A parallel runtime framework for communication intensive stream applications

### 2013-03-21 paper.tex
david's morning changes

Applications that follow streaming model~\cite{Thies:2002:SLS:647478.727935,Halbwachs91thesynchronous,Stephens95asurvey,muthukrishnan2005data} are expressed as a set of filters that operate on one or more tokens of data. The flow of these data tokens through the filters is referred to as a \textit{stream}. Since the data is accessed sequentially in the form of a stream the memory requirement to process the data is considerably less compared to the size of the enitre stream, thereby making it possible to operate on very large data sizes. Typical applications based on this model operate on data ranges in the order of terabytes or even bigger. Applications such as financial trading systems~\cite{agarwal2009faster}, database systems~\cite{Golab03datastream}, network data analysis~\cite{Gilbert01quicksand:quick,Sullivan:1998:TSM:1268256.1268258} are few examples that follow the streaming model and operate on very large data sets.

Traditional stream applications target multicore processors for performing their computations in parallel. As multicore processors grow in size, so do stream applications, processing more streams of data in parallel. However, in such cases the applications are typically limited by the traditional communication systems~\cite{Wu:2007:PPB:1323954.1323957,Wu:2007:PAL:1227865.1228021,Dobrescu09routebricks:exploiting,Han:2010:PGS:1851275.1851207,Kohler2000}. The reason being these systems have limited support for operating on parallel streams and are burdened by abstraction layers that provide compatibility for non-streaming applications.

>>>>

Applications that follow streaming model~\cite{Thies:2002:SLS:647478.727935,Halbwachs91thesynchronous,Stephens95asurvey,muthukrishnan2005data}
are expressed as a network of \textit{filters} connected by communication
\textit{channels}. Tokens of data flow along the communication channels
between filters. Each filter consumes on one or more \textit{streams}
of data from its input channels, and produces streams of data on its
ouput channels. For applications that fit the model well, stream processing
can greatly reduce the working memory required. Since the data flows through
a sequence of filters in a pipelined style, it is often possible to
operate on a relatively small window of the data at any given time,
thereby making it possible to operate on conceptually infinite
data streams.

Traditionally, the stream processing model was used primarily
signal processing type applications, where the application
processes a continuous stream of signal inputs. For example,
many video processing applications are modelled as a set of
image filters operating on a stream of video frame, where the
entire video may be hundreds of gigabytes in size, but the
processing can be done with just a few frames at a time.
However, in more recent years it has been found that the stream
model is highly suitable for many applications that process
very large data sets, in the order of terabytes or even larger.
These include applications such as financial trading
systems~\cite{agarwal2009faster},
database systems~\cite{Golab03datastream},
network data analysis~\cite{Gilbert01quicksand:quick,Sullivan:1998:TSM:1268256.1268258} are few examples that follow the streaming model and operate on very large data sets.

Stream processing is highly suited to multicore and parallel
computing, because the filters can typically execute in parallel.
Smaller stream applications often achieve high levels of parallelism
on multicore machines.  However, as the stream processing model
is used to address larger data processing problems, more computational
resources are needed, and a common solution is to distribute the
filters across a number of machines.
In such cases the applications are typically limited by the
traditional communication
~\cite{Wu:2007:PPB:1323954.1323957,Wu:2007:PAL:1227865.1228021,Dobrescu09routebricks:exploiting,Han:2010:PGS:1851275.1851207,Kohler2000}.
The reason being these systems have limited support for operating on
parallel streams and are burdened by abstraction layers that provide
compatibility for non-streaming applications.
